### 2021-10-29 Manuscript_draft.Rmd
Update Manuscript_draft.Rmd

---
title: "Occurrence and transmission potential of asymptomatic and pre-symptomatic SARS-CoV-2 infections: a living systematic review and meta-analysis"
author: "Diana Buitrago-Garcia, Aziz Mert Ipekci, Leonie Heron, Stefanie Hossmann, Hira Imeri, Georgia Salanti, Nicola Low"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output: html_document

#always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

#download packages
library(meta)
library(readxl)
library(tidyverse)
library(httr) # use to retrieve data from REDCap
library(kableExtra)
library(flextable)
library(dplyr)
library(RCurl)
library(tidyr)
library(ggplot2)
library(metafor)

#download data
# get the data directly from redcap:
# report #155 is Q1:
url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData <- list("token"=token,
                 content='report',
                 format='csv',
                 report_id='155',
                 csvDelimiter='',
                 rawOrLabel='raw',
                 rawOrLabelHeaders='raw',
                 exportCheckboxLabel='false',
                 returnFormat='csv'
)
response <- httr::POST(url, body = formData, encode = "form")

##########################################
# Q1 forest plot
##########################################

asymptomaticQ1 <- httr::content(response)

settings=c("Contact investigation",
           "Contact investigation, aggregated",
           "Outbreak investigation",
           "Statistical model",
           "Screening",
           "Hospitalised adults",
           "Hospitalised children",
           "Hospitalised adults & children",
           "Screening: institutional setting",
           "Screening: community setting",
           "Screening: occupational")

asymptomaticQ1$setting2<-factor(asymptomaticQ1$setting2, levels=1:11, labels=settings)
#combine contact investigation and contact investigation aggregated
asymptomaticQ1$setting2[asymptomaticQ1$setting2 == "Contact investigation, aggregated"] <- "Contact investigation"

# minor cleaning
asymptomaticQ1[asymptomaticQ1==9999]=NA

asymptomaticQ1$setting = asymptomaticQ1$setting2

data_long1 <- gather(asymptomaticQ1, cluster, total, c(q1_c1_total,q1_c2_total,q1_c3_total), factor_key=TRUE) %>% 
  mutate(id=1:nrow(.)) %>%
  select(record_id, author_1, setting, total, id, cluster)
data_long2 <-gather(asymptomaticQ1, cluster, events, c(q1_c1_event,q1_c2_event,q1_c3_event), factor_key=TRUE) %>% 
  mutate(id=1:nrow(.)) %>%
  select(events, id)

data_Q1 = merge(data_long1, data_long2, by="id")
data_Q1 = data_Q1[!is.na(data_Q1$total),]
data_Q1$cluster=factor(data_Q1$cluster,labels=c("1","2","3"),levels=c("q1_c1_total","q1_c2_total","q1_c3_total"))

data_Q1[data_Q1$record_id %in% asymptomaticQ1[is.na(asymptomaticQ1$q1_c2_event),]$record_id,]$cluster=NA

asymptomaticQ1=data_Q1

asymptomaticQ1$label=paste0("#", asymptomaticQ1$record_id, " ", asymptomaticQ1$author_1, " (2020)",ifelse(!is.na(asymptomaticQ1$cluster),paste0(" cluster:",asymptomaticQ1$cluster),"")) # [FU: ",asymptomaticQ1$fup_median,"]")

asymptomaticQ1$label=paste0("#", asymptomaticQ1$record_id, " ", asymptomaticQ1$author_1,ifelse(!is.na(asymptomaticQ1$cluster),paste0(" [cluster:",asymptomaticQ1$cluster,"]"),"")) # [FU: ",asymptomaticQ1$fup_median,"]")

#change clusters to descriptions
asymptomaticQ1$label[asymptomaticQ1$label == "#4479 Harada S [cluster:1]"] <- "#4479 Harada S [Patients]"
asymptomaticQ1$label[asymptomaticQ1$label == "#4479 Harada S [cluster:2]"] <- "#4479 Harada S [Healthcare workers]"
asymptomaticQ1$label[asymptomaticQ1$label == "#2892 Kennelly SP [cluster:1]"] <- "#2892 Kennelly SP [Nursing home residents]"
asymptomaticQ1$label[asymptomaticQ1$label == "#2892 Kennelly SP [cluster:2]"] <- "#2892 Kennelly SP [Nursing home staff]"
asymptomaticQ1$label[asymptomaticQ1$label == "#4968 van Buul LW [cluster:1]"] <- "#4968 van Buul LW [Nursing home residents]"
asymptomaticQ1$label[asymptomaticQ1$label == "#4968 van Buul LW [cluster:2]"] <- "#4968 van Buul LW [Healthcare workers]"
asymptomaticQ1$label[asymptomaticQ1$label == "#9442 Theuring S [cluster:1]"] <- "#9442 Theuring S [School students and staff]"
asymptomaticQ1$label[asymptomaticQ1$label == "#9442 Theuring S [cluster:2]"] <- "#9442 Theuring S [Household members]"
asymptomaticQ1$label[asymptomaticQ1$label == "#7465 van den Besselaar JH [cluster:1]"] <- "#7465 van den Besselaar JH [Healthcare workers]"
asymptomaticQ1$label[asymptomaticQ1$label == "#7465 van den Besselaar JH [cluster:2]"] <- "#7465 van den Besselaar JH [Residents]"
asymptomaticQ1$label[asymptomaticQ1$label == "#2826 Taylor J [cluster:1]"] <- "#2826 Taylor J [Healthcare personnel]"
asymptomaticQ1$label[asymptomaticQ1$label == "#2826 Taylor J [cluster:2]"] <- "#2826 Taylor J [Residents]"
asymptomaticQ1$label[asymptomaticQ1$label == "#3719 Ladhani SN [cluster:1]"] <- "#3719 Ladhani SN [Residents]"
asymptomaticQ1$label[asymptomaticQ1$label == "#3719 Ladhani SN [cluster:2]"] <- "#3719 Ladhani SN [Healthcare workers]"
#asymptomaticQ1$label[asymptomaticQ1$label == "#5551 Vohra [cluster:1]"] <- "#5551 Vohra [Presurgical patients]"
#asymptomaticQ1$label[asymptomaticQ1$label == "#5551 Vohra [cluster:2]"] <- "#5551 Vohra [Undergoing chemotherapy]"
#asymptomaticQ1$label[asymptomaticQ1$label == "#6526 Garibaldi [cluster:1]"] <- "#6526 Garibaldi [Residents]"
#asymptomaticQ1$label[asymptomaticQ1$label == "#6526 Garibaldi [cluster:2]"] <- "#6526 Garibaldi [Staff]"

data=asymptomaticQ1[order(asymptomaticQ1$setting,1/(1/asymptomaticQ1$events+1/(asymptomaticQ1$total-asymptomaticQ1$events))),]

data[is.na(data$setting),]$record_id
data=data[!is.na(data$setting),]

#cut data at end of jan 31
published_preprints <-c(5565,6219, 6685, 7030, 7465, 8249, 9442, 9484)
data <- data %>%
  filter(record_id <= 5296 | record_id %in% published_preprints)

asym_plot<-metaprop(events,total,data=data,sm = "PLOGIT", studlab=label, 
                    byvar=setting,# tau.common =TRUE,
                    prediction = TRUE,
                    control=list(stepadj=0.05, maxiter=10000))#, method ="INV") #, verbose=TRUE, digits=5, control=list(stepadj=0.5))

asym_plot

png(file = 'forest_meta_Q1.png',width=28,height=70, res=600, units="cm")

forest(asym_plot,  col.square=data$setting, sortvar = 1/seTE,
       #squaresize = 1/(1/data$events+1/(data$total-data$events)),
       #sortvar = 1/(1/events+1/(total-events)), 
       subgroup=TRUE, 
       just.studlab="left", colgap.studlab="2cm",
       #leftcols = c("label", "setting", "age", "design", "asym_denom"),
       #leftcols=c("studlab", "record_id", "event", "n"),
       xlab = "",just="left",comb.random=TRUE,test.subgroup.random=FALSE,
       comb.fixed=FALSE,test.subgroup.fixed=FALSE, predict=T)#, study.results=TRUE,test.subgroup=TRUE)
dev.off()

#sort by standard error
#read meta book and figure it out
#same as precision plot
pdf(file = 'forest_meta_Q1.pdf', width = 10, height = 28)

forest(asym_plot,  col.square=data$setting,sortvar = 1/(sqrt(seTE)),
       #squaresize = 1/(1/data$events+1/(data$total-data$events)),
       #sortvar = 1/(1/events+1/(total-events)), 
       subgroup=TRUE, 
       just.studlab="left", colgap.studlab="2cm",
       #leftcols = c("label", "setting", "age", "design", "asym_denom"),
       #leftcols=c("studlab", "record_id", "event", "n"),
       xlab = "",just="left",comb.random=TRUE,test.subgroup.random=FALSE,
       comb.fixed=FALSE,test.subgroup.fixed=FALSE, predict=T)#, study.results=TRUE,test.subgroup=TRUE)
dev.off()

q1_nstudies <- nrow(asym_plot$data) #total number of studies included in q1
q1_nppl <- prettyNum(sum(asymptomaticQ1$total), big.mark = ",") #total number of people included in q1
q1_nasymp <-  prettyNum(sum(asymptomaticQ1$events), big.mark = ",") #number of people asymptomatic included in q1

##########################################
# Q2.1 forest plot
##########################################

# get the data directly from redcap: 
# report #172 is Q2.1 SAR:
url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData2_1 <- list("token"=token,
                    content='report',
                    format='csv',
                    report_id='172',
                    csvDelimiter='',
                    rawOrLabel='raw',
                    rawOrLabelHeaders='raw',
                    exportCheckboxLabel='false',
                    returnFormat='json'
)
response2_1 <- httr::POST(url, body = formData2_1, encode = "form")
asymptomaticQ2_1 <- httr::content(response2_1)

#clean data
asymptomaticQ2_1[asymptomaticQ2_1=="9999;9999"]=NA #indicate as missing
asymptomaticQ2_1[asymptomaticQ2_1=="9999"]=NA #indicate as missing

#added for now
asymptomaticQ2_1 <- asymptomaticQ2_1[asymptomaticQ2_1$record_id < 5296, ]

asymptomaticQ2_1=asymptomaticQ2_1%>% #separate symp SAR into 2 variables
  separate(q3_sar_s, c("Ec","Nc"), ";", remove=FALSE) 
asymptomaticQ2_1=asymptomaticQ2_1%>% #separate asymp SAR into 2 variables
  separate(q3_sar_a, c("Ee_a","Ne_a"), ";", remove=FALSE) 
asymptomaticQ2_1=asymptomaticQ2_1%>% #separate presymp SAR into 2 variables
  separate(q3_sar_p, c("Ee_p","Ne_p"), ";", remove=FALSE)

asymptomaticQ2_1=asymptomaticQ2_1 %>% #change values to numeric
  mutate(Ec=as.numeric(Ec),
         Nc=as.numeric(Nc),
         Ee_a=as.numeric(Ee_a),
         Ne_a=as.numeric(Ne_a),
         Ee_p=as.numeric(Ee_p),
         Ne_p=as.numeric(Ne_p))

#create new df with asymp/presymp as subgroups
q2_1_df_a <- asymptomaticQ2_1 %>%
  select("record_id", "author_1", "Ee_a", "Ne_a", "Ec", "Nc", "region", "year") %>%
  rename(Ee = Ee_a,
         Ne = Ne_a) %>%
  filter(!is.na(Ee)) %>%
  mutate(group = "Asymptomatic")
q2_1_df_p <- asymptomaticQ2_1 %>%
  select("record_id", "author_1", "Ee_p", "Ne_p", "Ec", "Nc", "region", "year") %>%
  rename(Ee = Ee_p,
         Ne = Ne_p) %>%
  filter(!is.na(Ee)) %>%
  mutate(group = "Presymptomatic")
#bind rows
q2_1_df <- bind_rows(q2_1_df_a, q2_1_df_p)

#cut data at end of jan 31
published_preprints <-c(5565,6219, 6685, 7030, 7465, 8249, 9442, 9484)
q2_1_df <- q2_1_df %>%
  filter(record_id <= 5296 | record_id %in% published_preprints)

#metaanalysis for asymptomatic transmission

q2_1_asymp <- metabin(event.e = Ee,
                      n.e = Ne,
                      event.c = Ec,
                      n.c = Nc, 
                      data = q2_1_df,
                      studlab = author_1,
                      prediction = TRUE,
                      sm = "RR",
                      method = "MH",
                      MH.exact = TRUE, 
                      byvar = group,
                      comb.random = TRUE,
                      comb.fixed = FALSE, 
                      label.e = "",
                      label.c = "Symptomatic")
q2_1_asymp

pdf(file = 'forest_meta_Q2_1.pdf', width = 8, height = 6)

forest(q2_1_asymp, sortvar = (1/(sqrt(seTE))), overall = FALSE,
       overall.hetstat = FALSE,
       leftlabs = c(NA, "      E", "      N", "            E", "        N"))

pdf(file = 'forest_meta_Q2_1.pdf', width = 10, height = 6)

forest(q2_1_asymp, sortvar = (1/(sqrt(seTE))), overall = FALSE,
       overall.hetstat = FALSE,
       leftlabs = c(NA, "      E", "      N", "            E", "        N"))

dev.off()

q2_1_nstudies <- nrow(q2_1_asymp$data) #total number of studies included in q1

##########################################
# supplementary information - by precision
##########################################

#metaanalysis for asymptomatic transmission

q2_1_asymp_precision <- metabin(event.e = Ee,
                                n.e = Ne,
                                event.c = Ec,
                                n.c = Nc, 
                                data = q2_1_df,
                                studlab = author_1,
                                sm = "RR",
                                byvar = group)

png(file = 'forest_meta_Q2_1.png', width = 600, height = 300)

forest(q2_1_asymp, sortvar = (1/(sqrt(seTE))), overall = FALSE,
       overall.hetstat = FALSE,
       leftlabs = c(NA, "      E", "      N", "            E", "        N"))

dev.off()

##########################################
# table on characteristics of Q2.1 studies
##########################################

#table not necessary - all studies in q2.1 described in full table on q1, and numbers are in forest plot

#get data
#Q2_1_table <- httr::content(response2_1)

#clean countries
#Q2_1_table <- Q2_1_table %>%
 # mutate(country = ifelse(country == 10, "Australia", country),
  #       country = ifelse(country == 11, "Austria", country),
  #       country = ifelse(country == 15, "Bahrain", country),
  #       country = ifelse(country == 19, "Belgium", country),
  #       country = ifelse(country == 28, "Brunei", country),
  #       country = ifelse(country == 34, "Canada", country),
  #       country = ifelse(country == 40, "China", country),
  #       country = ifelse(country == 41, "Colombia", country),
  #       country = ifelse(country == 51, "Denmark", country),
  #       country = ifelse(country == 62, "France", country),
  #       country = ifelse(country == 63, "French Guyana", country),
  #       country = ifelse(country == 67, "Germany", country),
  #       country = ifelse(country == 69, "Greece", country),
  #       country = ifelse(country == 78, "Iceland", country),
  #       country = ifelse(country == 79, "India", country),
  #       country = ifelse(country == 83, "Ireland", country),
  #       country = ifelse(country == 85, "Italy", country),
  #       country = ifelse(country == 87, "Japan", country),
  #       country = ifelse(country == 92, "Kuwait", country),
  #       country = ifelse(country == 105, "Malaysia", country),
  #       country = ifelse(country == 121, "Netherlands", country),
  #       country = ifelse(country == 128, "Norway", country),
  #       country = ifelse(country == 888, "Other", country),
  #       country = ifelse(country == 138, "Portugal", country),
  #       country = ifelse(country == 146, "Saudi Arabia", country),
  #       country = ifelse(country == 157, "South Korea", country),
  #       country = ifelse(country == 160, "Spain", country),
  #       country = ifelse(country == 174, "Turkey", country),
  #       country = ifelse(country == 176, "Uganda", country),
  #       country = ifelse(country == 180, "United Kingdom", country),
  #       country = ifelse(country == 181, "United States of America", country),
  #       country = ifelse(country == 185, "Vietnam", country))
#clean regions

#Q2_1_table$region[Q2_1_table$region == "none" | Q2_1_table$region == "None."] <- NA
#create location variable
#Q2_1_table$location <- Q2_1_table$country
#Q2_1_table <- Q2_1_table %>%
#  mutate(location = ifelse(!is.na(region), paste0(country, ", ", region), country))
#drop country and region variables
#Q2_1_table$country <- NULL
#Q2_1_table$region <- NULL

#author column
#Q2_1_table$study <- paste0(Q2_1_table$author_1, " (", Q2_1_table$year, ")")
#Q2_1_table$author_1 <- NULL
#Q2_1_table$record_id <- NULL
#Q2_1_table$year <- NULL

#clean settings
#Q2_1_table$setting <- "Not available"
#Q2_1_table <- Q2_1_table %>%
 # mutate(setting = ifelse(setting2 == 1, "Contact investigation", setting),
#         setting = ifelse(setting2 == 2, "Contact investigation, aggregated", setting),
#         setting = ifelse(setting2 == 3, "Outbreak investigation", setting),
#         setting = ifelse(setting2 == 4, "Statistical model", setting),
#         setting = ifelse(setting2 == 5, "Screening", setting),
#         setting = ifelse(setting2 == 6, "Hospitalised adults", setting),
#         setting = ifelse(setting2 == 7, "Hospitalised children", setting),
#         setting = ifelse(setting2 == 8, "Hospitalised children and adults", setting),
#         setting = ifelse(setting2 == 9, "Screening: hospitalised patients", setting),
#         setting = ifelse(setting2 == 10, "Screening: general population", setting),
#         setting = ifelse(setting2 == 11, "Screening: occupational", setting))
#Q2_1_table$setting2 <- NULL

#reorder columns
#Q2_1_table <- data.frame(Q2_1_table$study,
 #                        Q2_1_table$setting,
#                         Q2_1_table$location)

#names(Q2_1_table) <- c("study", "setting", "location")

###############################################################
# Question 2.2: results output ################################
# Proportion of transmission from asymp or pre-symp ###########
###############################################################

# get the data directly from REDCap: 
# report #157 is Q2.2:

url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData2_2 <- list("token"=token,
                    content='report',
                    format='csv',
                    report_id='157',
                    csvDelimiter='',
                    rawOrLabel='raw',
                    rawOrLabelHeaders='raw',
                    exportCheckboxLabel='false',
                    returnFormat='csv'
)
response2_2 <- httr::POST(url, body = formData2_2, encode = "form")
asymptomaticQ2_2 <- httr::content(response2_2)

############Old Codes###############
#clean data
# asymptomaticQ2_2[asymptomaticQ2_2=="9999;9999"]=NA #indicate as missing
# asymptomaticQ2_2[asymptomaticQ2_2 == "9999"] = NA #indicate as missing
# 
# #create a separate dataset for the records with >1 result
# two_models <- asymptomaticQ2_2
# two_models$author_1 <- gsub(",","",two_models$author_1)
# #separate into 2 variables
# two_models = two_models %>% 
#   separate(q3_pp_m, c("q3_pp_m_1","q3_pp_m_2"), ";")
# two_models = two_models %>% 
#   separate(q3_pp_l, c("q3_pp_l_1","q3_pp_l_2"), ";") 
# two_models = two_models %>%
#   separate(q3_pp_u, c("q3_pp_u_1","q3_pp_u_2"), ";")
# two_models = two_models %>% 
#   separate(q3_setting, c("q3_setting_1","q3_setting_2"), ";") 
# 
# two_models <- two_models %>% #change values to numeric
#   mutate(q3_pp_m_1=as.numeric(q3_pp_m_1),
#          q3_pp_m_2=as.numeric(q3_pp_m_2),
#          q3_pp_l_1=as.numeric(q3_pp_l_1),
#          q3_pp_l_2=as.numeric(q3_pp_l_2),
#          q3_pp_u_1=as.numeric(q3_pp_u_1),
#          q3_pp_u_2=as.numeric(q3_pp_u_2))
# 
# 
# two_models <- two_models[!is.na(two_models$q3_pp_m_2),]
# 
# substrRight <- function(x, n){
#   substr(x, nchar(x)-n+1, nchar(x))
# }
# 
# L1 <- two_models %>%
#   pivot_longer(cols = c(6:7),
#                names_to = "model",
#                values_to = "q3_pp_m") 
# 
# L1$label <- "Value"
# for (i in 1:nrow(L1)) {
#   if (L1$model[i] == "q3_pp_m_1") {
#   L1$label[i] <- paste(L1$author_1[i], " [", L1$q3_setting_1[i], "]", sep = "")
#   }else{
#     L1$label[i] <- paste(L1$author_1[i], " [", L1$q3_setting_2[i], "]", sep = "")
#   }
#   
# }
# 
# L1$label2 <- paste(L1$author_1, " [", substrRight(L1$model, 1), "]", sep = "")
# L2 <- two_models %>%
#   pivot_longer(cols = c(8:9),
#                names_to = "model",
#                values_to = "q3_pp_l") %>%
#   select("record_id", "model", "q3_pp_l", "author_1", "q3_setting_1", "q3_setting_2")
# 
# L2$label2 <- paste(L2$author_1, " [", substrRight(L2$model, 1), "]", sep = "")
# L2$label <- "Value"
# i = 1
# for (i in 1:nrow(L2)) {
#   if (L2$model[i] == "q3_pp_l_1") {
#     sett -> L2$q3_setting_1[i]
#   }else{
#     sett -> L2$q3_setting_2[i]
#   }
#   L2$label[i] <- paste(L2$author_1[i], " [", sett, "]", sep = "")
# }
# 
# i=1
# for (i in 1:nrow(L2)) {
#   if (L2$model[i] == "q3_pp_l_1") {
#     L2$label[i] <- paste(L2$author_1[i], " [", L2$q3_setting_1[i], "]", sep = "")
#   }else{
#     L2$label[i] <- paste(L2$author_1[i], " [", L2$q3_setting_2[i], "]", sep = "")
#   }
#   
# }
# 
# L2 <- L2 %>% select("q3_pp_l", "label")
# L3 <- two_models %>%
#   pivot_longer(cols = c(10:11),
#                names_to = "model",
#                values_to = "q3_pp_u") %>%
#   select("record_id", "model", "q3_pp_u", "author_1", "q3_setting_1", "q3_setting_2")
# 
# L3$label <- "Value"
# i = 1
# for (i in 1:nrow(L3)) {
#   if (L3$model[i] == "q3_pp_u_1") {
#     L3$label[i] <- paste(L3$author_1[i], " [", L3$q3_setting_1[i], "]", sep = "")
#   }else{
#     L3$label[i] <- paste(L3$author_1[i], " [", L3$q3_setting_2[i], "]", sep = "")
#   }
#   
# }
# 
# L3$label2 <- paste(L3$author_1, " [", substrRight(L3$model, 1), "]", sep = "")
# 
# L3 <- L3 %>% select("q3_pp_u", "label")
# L4 <- two_models %>%
#   pivot_longer(cols = c(13:14),
#                names_to = "model",
#                values_to = "setting") %>%
#   select("record_id", "model", "setting", "author_1")
# L4$label2 <- paste(L4$author_1, " [", substrRight(L4$model, 1), "]", sep = "")
# L4$label <- paste(L4$author_1, " [", L4$setting, "]", sep = "")
# 
# L4 <- L4 %>% select("setting", "label")
# #merge together
# Lall <- merge(L2, L3, by = "label")   
# Lall <- merge(Lall, L4, by = "label") 
# Lall <- merge(Lall, L1, by = "label")
# Lall <- Lall %>%
#   select("record_id", "author_1", "q3_pa_m", "q3_pa_l", "q3_pa_u", 
#          "q3_pp_m", "q3_pp_l", "q3_pp_u", "question_3_complete", "setting", 
#          "label")
# 
# #clean original dataset
# asymptomaticQ2_2$label <- asymptomaticQ2_2$author_1
# asymptomaticQ2_2 <- asymptomaticQ2_2 %>%
#   mutate(setting = q3_setting)
# asymptomaticQ2_2$q3_setting <- NULL
# asymptomaticQ2_2 <- as.data.frame(asymptomaticQ2_2)
# attr(asymptomaticQ2_2, 'spec') <- NULL
# str(asymptomaticQ2_2)
# asymptomaticQ2_2$q3_pp_m <- as.numeric(asymptomaticQ2_2$q3_pp_m)
# asymptomaticQ2_2$q3_pp_l <- as.numeric(asymptomaticQ2_2$q3_pp_l)
# asymptomaticQ2_2$q3_pp_u <- as.numeric(asymptomaticQ2_2$q3_pp_u)
# 
# #add extra rows to original dataset
# Q2_2 <- bind_rows(asymptomaticQ2_2, Lall)
# #if q3_pp_m and q3_pa_m both NA then delete
# Q2_2 <- Q2_2[!is.na(Q2_2$q3_pa_m) | !is.na(Q2_2$q3_pp_m),]
# 
# #create new variable for all estimates and indicate asymp/presymp with new variable
# #fill Q2_2 first with estimates from q3_pp_m then q3_pa_m
# Q2_2 <- Q2_2 %>%
#   mutate(estimate = ifelse(!is.na(q3_pp_m),
#                            q3_pp_m,
#                            q3_pa_m),
#          lower_bound = ifelse(!is.na(q3_pp_m),
#                                  q3_pp_l,
#                                  q3_pa_l),
#          upper_bound = ifelse(!is.na(q3_pp_m),
#                                  q3_pp_u,
#                                  q3_pa_u),
#          subgroup = ifelse(!is.na(q3_pp_m),
#                            "presymp",
#                            "asymp"))
# #some records have presymp estimates in estimate now - must make new observation
# both <- Q2_2 %>%
#   filter(!is.na(Q2_2$q3_pa_m) & !is.na(Q2_2$q3_pp_m)) %>%
#   mutate(estimate = q3_pa_m,
#          lower_bound = q3_pa_l,
#          upper_bound = q3_pa_u,
#          subgroup = "asymp")
# Q2_2 <- bind_rows(Q2_2, both)
# #remove unnecessary cols
# Q2_2 <- Q2_2 %>%
#   select(c("record_id", "label", "setting", "estimate", "lower_bound", "upper_bound", "subgroup"))
# #round numbers to 2 decimal points
# Q2_2$estimate <- round(Q2_2$estimate, digits = 2)
# Q2_2$lower_bound <- round(Q2_2$lower_bound, digits = 2)
# Q2_2$upper_bound <- round(Q2_2$upper_bound, digits = 2)
# 
# ###############################################################
# # forest plot
# 
# 
# #prepare data
# Q2_2_df <- Q2_2
# #create label for proportions
# Q2_2_df$prop <- sprintf("%2.2f", Q2_2_df$estimate)
# #create label for 95% CIs
# Q2_2_df$CI_lower <- sprintf("%2.2f", Q2_2_df$lower_bound)
# Q2_2_df$CI_upper <- sprintf("%2.2f", Q2_2_df$upper_bound)
# Q2_2_df$CIs <- paste("[", Q2_2_df$CI_lower, #paste together 95% CIs
#                      ";", Q2_2_df$CI_upper, "]", 
#                      sep = "")
# Q2_2_df <- Q2_2_df %>%
#   mutate(CI = ifelse(is.na(Q2_2_df$CI_lower), NA, CIs))
# Q2_2_df$CI[Q2_2_df$CI == "[NA;NA]"] <- NA
# str(Q2_2_df)
# #col to indicate headings/studies in plot
# Q2_2_df$type <- 2
# #new row for asymptomatic label in plot
# Asymp_label <- data.frame(NA, "Asymptomatic transmission", NA, NA, NA, NA, NA, 
#                           NA, NA, NA, NA, NA,1)
# names(Asymp_label) <- names(Q2_2_df)
# #new row for asymptomatic label in plot
# Presymp_label <- data.frame(NA, "Presymptomatic transmission", NA, NA, NA, NA, NA, 
#                             NA, NA, NA, NA, NA, 1)
# names(Presymp_label) <- names(Q2_2_df)
# #add to exisiting dataframe
# Q2_2_df <- rbind(Q2_2_df, Asymp_label, Presymp_label)
# #font indicator
# Q2_2_df$fontface = ifelse(Q2_2_df$type == 1, "italic","plain")
# str(Q2_2_df$fontface)
# 
# #arrange studies by subgroup
# asymp_n <- sum(Q2_2_df$subgroup == "asymp", #number of asymp studies
#                na.rm = TRUE) 
# presymp_n <- sum(Q2_2_df$subgroup == "presymp", #number of presymp studies
#                  na.rm = TRUE) 
# Q2_2_df$line[Q2_2_df$label == "Asymptomatic transmission"] <- 1
# #new row for space in plot
# space <- data.frame(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA)
# names(space) <- names(Q2_2_df)
# #add to exisiting dataframe
# Q2_2_df <- rbind(Q2_2_df, space)
# Q2_2_df$line[is.na(Q2_2_df$label)] <- 2 + asymp_n
# Q2_2_df$line[Q2_2_df$label == "Presymptomatic transmission"] <- 3 + asymp_n
# Q2_2_df_a <- Q2_2_df %>% #order the asymp studies
#   filter(subgroup == "asymp") %>%
#   mutate(line = 2:(asymp_n + 1))
# Q2_2_df_p <- Q2_2_df %>% #order the presymp studies
#   filter(subgroup == "presymp") %>%
#   mutate(line = (asymp_n + 4):(presymp_n + asymp_n + 3))
# Q2_2_df <- Q2_2_df[is.na(Q2_2_df$record_id),]
# #join all rows again
# Q2_2_df <- rbind(Q2_2_df, Q2_2_df_a, Q2_2_df_p)
# 
# #create forest plot
# p <- ggplot() + 
#   geom_point(data = Q2_2_df, #add estimates 
#              aes(y = line, 
#                  x = estimate),
#              fill = "gray", 
#              shape = 22, 
#              color = "black",
#              size = 4) +
#   geom_errorbarh(data = Q2_2_df, #add error bars 
#                  aes(y = line, 
#                      xmin = lower_bound, 
#                      xmax = upper_bound), 
#                  height=0.3) +
#   geom_text(aes(y = c(-1, -1, -1),  #add cols for study, estimates and CIs
#                 x = c(-0.2, 1.05, 1.12), 
#                 label=c("Study", "Prop.", "95% CI")), 
#             hjust = 0, 
#             fontface = "bold") +
#   geom_text(data = Q2_2_df, #add estimate labels
#             aes(y = line, 
#                 x = 1.05, 
#                 label = prop), 
#             hjust = 0) +
#   geom_text(data = Q2_2_df, #add CI labels
#             aes(y = line,
#                 x = 1.12,
#                 label = CI), 
#             hjust = 0) +
#   geom_text(data = Q2_2_df, #add study labels
#             aes(y= line, 
#                 x=-0.2, 
#                 label = label,
#                 hjust = 0,
#                 fontface = fontface)) +
#   theme_void() + #remove background
#   scale_y_reverse() +
#   geom_segment(data = Q2_2_df, #add line below point estimates
#                aes(y = max(line) + 1, x = 0, xend = 1, 
#                    yend = max(line) + 1)) +
#   geom_segment(data = Q2_2_df, #add dash for 0
#                aes(y = max(line) + 1, yend = max(line) + 1.3, 
#                    x = 0, xend = 0)) +
#   geom_segment(data = Q2_2_df, #add dash for 0.25
#                aes(y = max(line) + 1, yend = max(line) + 1.3,
#                    x = 0.25, xend = 0.25)) +
#   geom_segment(data = Q2_2_df, #add dash for 0.5
#                aes(y = max(line)+1, yend=max(line) + 1.3,
#                    x=0.5, xend=0.5)) +
#   geom_segment(data = Q2_2_df, #add dash for 0.75
#                aes(y = max(line) + 1, yend = max(line) + 1.3,
#                    x=0.75, xend = 0.75)) +
#   geom_segment(data = Q2_2_df, #add dash for 1
#                aes(y = max(line) + 1, 
#                    yend = max(line) + 1.3,
#                    x = 1, xend = 1)) +
#   geom_text(aes(y = rep(max(Q2_2_df$line) + 2,5), #add numbers to line
#                 x = 0:4/4,
#                 label = c(0, 0.25, 0.5, 0.75, 1))) +
#   scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.5))

p

#create image file
png(file="forest_Q2_2.png", 
    res=300, 
    height=35, width=42, units="cm")
p
dev.off()

##########################################
# table on characteristics of Q2.2 studies
##########################################

#get data
Q2_2_table <- httr::content(response2_2)

#author column
Q2_2_table$study <- paste0(Q2_2_table$author_1, " (", Q2_2_table$year, ")")
Q2_2_table$author_1 <- NULL
Q2_2_table$record_id <- NULL
Q2_2_table$year <- NULL

#subset df
Q2_2_table <- Q2_2_table %>%
  select(study, q3_setting)

#some pre-processing HI
Q2_2_table$q3_setting <- gsub(";([[:alpha:]])", "; \\1", Q2_2_table$q3_setting)
Q2_2_table$q3_setting <- gsub("/([[:alpha:]])", ", \\1", Q2_2_table$q3_setting)
Q2_2_table$q3_setting[15] <- gsub(",", ";",Q2_2_table$q3_setting[15])
q2_2_nstudies <- nrow(Q2_2_table) #total number of studies included in q1

```

 **We included a total of `r q1_nstudies` studies that reported empirical data about `r q1_nppl` people with SARS-CoV-2 infection (`r q1_nasymp` defined as having asymptomatic infection), `r q2_1_nstudies` studies for Q2.1, and `r q2_2_nstudies` studies for Q2.2.**

**We included `r q1_nstudies` studies that reported empirical data about `r q1_nppl` people with SARS-CoV-2 infection (`r q1_nasymp` defined as having asymptomatic infection) (Table 1).**

![Figure 1. Forest plot of proportion  of people with asymptomatic SARS-CoV-2 infection, stratified by setting and ordered by precision](forest_meta_Q1.png)

The main risks of bias across all categories of empirical studies were in the selection and enrolment of people with asymptomatic infection and mismeasurement of asymptomatic status because of absent or incomplete definitions (S3 Fig). Sources of bias specific to studies in particular settings are discussed with the relevant results.

```{r table2, ft.split=TRUE, tab.cap='Table 1. Characteristics of studies included for Q2.1', echo = FALSE}
#create table
# Q2_1_table <- Q2_1_table %>%
#   arrange(setting)
# 
# table2 <- as_grouped_data(Q2_1_table, groups = "setting")
# table2 <- flextable(table2)
# table2 <- table2 %>%
#   set_header_labels(setting = "Setting",
#                     location = "Country, region", 
#                     study = "Study") %>%
#   set_table_properties(layout = "autofit") %>%
#   theme_vanilla()
# table2 <- width(table2, width = 10)
# table2
# 
# Q2_1_table <- Q2_1_table %>%
#   arrange(setting)
# table2 <- as_grouped_data(Q2_1_table, groups = "setting")
# table2 %>%
#   as_flextable() %>%
#   set_header_labels(setting = "Setting",
#                     location = "Country, region", 
#                     study = "Study") %>%
#   set_table_properties(layout = "autofit") %>% fit_to_width(20) %>% 
#   theme_vanilla()

```

![Figure 2.1. Forest plot of secondary attack rate comparing infections in contacts of asymptomatic and presymptomatic index cases with infections in contacts of symptomatic cases, ordered by precision](forest_meta_Q2_1.png)
```{r table3, ft.split=TRUE, tab.cap='Table 1. Characteristics of studies included for Q2.2', echo = FALSE}
#create table
table2 <- flextable(Q2_2_table)

table2 %>% 
  set_header_labels(study = "Study",
                    q3_setting = "Setting") %>%
  set_table_properties(layout = "autofit") %>%
  theme_vanilla()

```

![Figure 4. Forest plot of proportion of SARS-CoV-2 infection resulting from asymptomatic or presymptomatic transmission](forest_Q2_2.png)