### 2021-11-03 LSR-Progress-Report.Rmd
rerun

---
title: "LSR Progress Report"
author: "L Heron"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE, error = TRUE}

library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(kableExtra)

###########################################
#verification set up code
###########################################

url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData <- list("token"=token,
                 content='report',
                 format='csv',
                 report_id='159',
                 csvDelimiter='',
                 rawOrLabel='raw',
                 rawOrLabelHeaders='raw',
                 exportCheckboxLabel='false',
                 returnFormat='csv'
)

response <- httr::POST(url, body = formData, encode = "form")
result <- httr::content(response)

ref_asymp2 = result %>% 
  mutate(status2=ifelse((is.na(screen1_veri) & screening_api_complete==1), "Ready to be assigned for verification", NA),
         status2=ifelse((!is.na(screen1_veri) & screening_api_complete==1), "Assigned for verification", status2),
         status2=ifelse((!is.na(screen1_veri) & screening_api_complete==0), "Verified: resolution needed", status2),
         status2=ifelse((!is.na(screen1_veri) & screening_api_complete==2), "Verified: Done", status2))

ref_asymp2 <- ref_asymp2[1874:length(ref_asymp2$record_id),]

tmpveri=ref_asymp2 %>%
  group_by(screen1_veri, status2) %>%
  summarize(n=n())

tmpveri = tmpveri[!is.na(tmpveri$status2), ]

# current number of verified records
verified <- sum(tmpveri[tmpveri$status2=="Verified: Done",]$n)
#total assigned for verification
assigned_veri <- sum(tmpveri[tmpveri$status2=="Assigned for verification",]$n)
#waiting to be assigned
unassigned_veri <- sum(tmpveri[tmpveri$status2=="Ready to be assigned for verification",]$n)
#resolution needed
resolve <- sum(tmpveri[tmpveri$status2=="Verified: resolution needed",]$n)

tmpveri[tmpveri$status2=="Assigned for verification",]
sum(tmpveri[tmpveri$status2=="Assigned for verification",]$n)
tmpveri[tmpveri$status2=="Verified: resolution needed",]
sum(tmpveri[tmpveri$status2=="Ready to be assigned for verification",]$n)
#total needing resolution:
sum(tmpveri[tmpveri$status2=="Verified: resolution needed",]$n)

###########################################
#verification plot
###########################################

veriplot<-ggplot(tmpveri,aes(screen1_veri, n, fill=status2))+
  geom_bar(stat="identity")+
  coord_flip()  +
  scale_fill_manual(values=c("tomato1", "forestgreen", "deepskyblue1", "orchid4")) + 
  xlab("Reviewers") + 
  ylab("Number of records")

###########################################
#screening set up code
###########################################

formData2 <- list("token"=token,
                  content='report',
                  format='csv',
                  report_id='159',
                  csvDelimiter='',
                  rawOrLabel='raw',
                  rawOrLabelHeaders='raw',
                  exportCheckboxLabel='false',
                  returnFormat='csv'
)

response2 <- httr::POST(url, body = formData2, encode = "form")
result2 <- httr::content(response2)

ref_asymp = result2 %>% 
  mutate(status=ifelse((!is.na(screen1_assigned) & !is.na(comment)),"Screening complete",NA),
         status=ifelse((!is.na(screen1_assigned) & is.na(comment)),"Assigned & incomplete",status),
         status=ifelse(is.na(screen1_assigned),"Unassigned",status))
ref_asymp <- ref_asymp[1874:length(ref_asymp$record_id),]

tmp2=ref_asymp %>%
  group_by(screen1_assigned, status) %>%
  summarize(n=n())

#screening plot 
screenplot <- ggplot(tmp2,aes(screen1_assigned, n, fill=status))+
  geom_bar(stat="identity")+ 
  coord_flip() + xlab("Reviewers") + ylab("Number of records")

#assigned and still to screen
assigned_screen <- sum(tmp2[tmp2$status=="Assigned & incomplete",]$n)

#screened up to now
screen_complete <- (sum(tmp2[tmp2$status=="Screening complete",]$n))

#total screened and need to be screened
totaltoscreen <- (sum(tmp2[tmp2$status=="Assigned & incomplete",]$n)+sum(tmp2[tmp2$status=="Screening complete",]$n))+sum(tmp2[tmp2$status=="Unassigned",]$n)
# percentage
fraction <- screen_complete/totaltoscreen
percent <- paste(round(100*fraction), "%", sep="")

###########################################
#overall progress set up code
###########################################

url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData <- list("token"=token,
                 content='report',
                 format='csv',
                 report_id='159',
                 csvDelimiter='',
                 rawOrLabel='raw',
                 rawOrLabelHeaders='raw',
                 exportCheckboxLabel='false',
                 returnFormat='csv'
)

response <- httr::POST(url, body = formData, encode = "form")
result <- httr::content(response)

ref_asymp <- result %>% filter(screen1_assigned!="MC_X")

ref_asymp1 = ref_asymp %>%
  mutate(status=ifelse(screen1_assigned=="","New records",NA),
         status=ifelse((screen1_assigned!="" & comment==""),"Assigned",status),
         status=ifelse(((screen1_veri=="" | screen1_veri!="") & screening_api_complete==1), "Screened but not verified", status),
         status=ifelse((screen1_veri!="" & review ==1 & screening_api_complete==2), "Included",status),
         status=ifelse((screen1_veri!="" & review ==2 & screening_api_complete==2), "Excluded", status))

ref_asymp1$status[is.na(ref_asymp1$status)] <- "Help needed"
ref_asymp1 = filter(ref_asymp1, !(status %in% "Help needed"))
tmp0=ref_asymp1 %>% group_by(status) %>%
  summarize(n=n())
pre=tmp0[tmp0$status=="Screened but not verified",]$n

ref_asymp1$status[ref_asymp1$status == "Assigned" | ref_asymp1$status == "Screened but not verified"] <- "Records in screening"
ref_asymp1$status <- factor(ref_asymp1$status, levels = c("Included","Excluded","Records in screening","New records"))

p1<-ggplot(ref_asymp1,aes(x=1,fill=factor(ref_asymp1$status)))+
  geom_bar()+
  coord_flip()+
  scale_fill_manual(name="",breaks=c("New records", "Records in screening","Excluded","Included"),values=c("orange","yellow2","steelblue3","green3"))+
  ylab("Number of citations")+
  guides(fill=guide_legend(ncol=2,nrow=2,byrow=TRUE)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        #panel.border = element_blank(),
        #panel.grid.major = element_blank(),
        #panel.grid.minor = element_blank(),
        legend.position = "bottom",legend.spacing.x = unit(0.5,"cm"),
        plot.margin = margin(5,.5,2,.4, "cm"))

tmp1=ref_asymp1 %>% group_by(status) %>%
  summarize(n=n())

#Assigned and incomplete:
#pub=tmp1[tmp1$status=="Assigned",]$n
include=tmp1[tmp1$status=="Included",]$n[1]
exclude=tmp1[tmp1$status=="Excluded",]$n[1]
#veri=include+exclude
screened=pre+include+exclude
screeninprogress=tmp1[tmp1$status=="Records in screening",]$n[1]

text=paste0("From June 10th until ", format(Sys.Date(),"%d.%m.%Y"), ",\nthe search strategy has\nretrieved ",length(ref_asymp$record_id)," results.\n\n* ",
            screeninprogress, " records in screening\n* ", exclude," excluded\n* ",
            include, " included")

#total complete and still to be screened
total <- sum(tmp1[tmp1$status=="Included",]$n)+sum(tmp1[tmp1$status=="Excluded",]$n)+sum(tmp1[tmp1$status=="Records in screening",]$n)
#number included
included <-sum(tmp1[tmp1$status=="Included",]$n)
#number excluded
excluded <- sum(tmp1[tmp1$status=="Excluded",]$n)
#number still in the screening process
screening <- total - (included + excluded)
#number yet to be assigned
unassigned <- sum(tmp2[tmp2$status=="Unassigned",]$n)

###progress table###

#The last record we will consider including in the latest update is record #5296 
#(cut off being 31st Jan (date records entered COAP database))
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(kableExtra)

url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData2 <- list("token"=token,
                  content='report',
                  format='csv',
                  report_id='159',
                  csvDelimiter='',
                  rawOrLabel='raw',
                  rawOrLabelHeaders='raw',
                  exportCheckboxLabel='false',
                  returnFormat='csv'
)
response2 <- httr::POST(url, body = formData2, encode = "form")
result2 <- httr::content(response2)

ref_asymp = result2 %>% 
  mutate(status=ifelse((!is.na(screen1_assigned) & !is.na(comment)),"Screening complete",NA),
         status=ifelse((!is.na(screen1_assigned) & is.na(comment)),"Assigned & incomplete",status),
         status=ifelse(is.na(screen1_assigned),"Unassigned",status))
ref_asymp <- ref_asymp[1874:length(ref_asymp$record_id),]
#31 Jan is #5296, 28 Feb is 5856, 31 Mar is 6535, 30 Apr is 7176, 31 May is 7713, 30 Jun is 8332, 31 Jul is 8959
ref_asymp_latest <- ref_asymp %>%
  filter(record_id < 7025) #filter to last record that would be included
until_Jan31 <- ref_asymp %>%
  filter(record_id <= 5296) #filter to last record that would be included
until_Feb28 <- ref_asymp %>%
  filter(record_id <= 5856) #filter to last record that would be included
until_Mar31 <- ref_asymp %>%
  filter(record_id <= 6535) #filter to last record that would be included
until_Apr30 <- ref_asymp %>%
  filter(record_id <= 7176) #filter to last record that would be included
until_May31 <- ref_asymp %>%
  filter(record_id <= 7713) #filter to last record that would be included
until_Jun30 <- ref_asymp %>%
  filter(record_id <= 8332) #filter to last record that would be included
until_Jul31 <- ref_asymp %>%
  filter(record_id <= 8959) #filter to last record that would be included
until_Aug31 <- ref_asymp %>%
  filter(record_id <= 9483) #filter to last record that would be included
until_Sep30 <- ref_asymp %>%
  filter(record_id <= 10074) #filter to last record that would be included
until_Oct05 <- ref_asymp %>%
  filter(record_id <= 10154) #filter to last record that would be included

#create df
progress <- data.frame(search_date = c("Jan 31", "Feb 28", "31 Mar", "30 Apr", "31 May", "30 Jun", "31 Jul", "31 Aug", "30 Sep", "5 Oct"),
                       redcap_id = c(5296, 5856, 6535, 7176, 7713, 8332, 8959, 9483, 10074, 10154),
                       included = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA),
                       screening = c(nrow(until_Jan31[until_Jan31$status == "Assigned & incomplete",]), 
                                     nrow(until_Feb28[until_Feb28$status == "Assigned & incomplete",]),
                                     nrow(until_Mar31[until_Mar31$status == "Assigned & incomplete",]),
                                     nrow(until_Apr30[until_Apr30$status == "Assigned & incomplete",]),
                                     nrow(until_May31[until_May31$status == "Assigned & incomplete",]),
                                     nrow(until_Jun30[until_Jun30$status == "Assigned & incomplete",]),
                                     nrow(until_Jul31[until_Jul31$status == "Assigned & incomplete",]),
                                     nrow(until_Aug31[until_Aug31$status == "Assigned & incomplete",]),
                                     nrow(until_Sep30[until_Sep30$status == "Assigned & incomplete",]),
                                     nrow(until_Oct05[until_Oct05$status == "Assigned & incomplete",])), 
                       verification = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA),
                       disagreements = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA),
                       data_extraction = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA))

#until_Jan31 <- ref_asymp %>%
# filter(record_id <= 5296) #filter to last record that would be included
#until_Feb28 <- ref_asymp %>%
#  filter(record_id <= 5856) #filter to last record that would be included
#until_Mar31 <- ref_asymp %>%
#  filter(record_id <= 6535) #filter to last record that would be included
#until_Apr30 <- ref_asymp %>%
#  filter(record_id <= 7176) #filter to last record that would be included
#until_May31 <- ref_asymp %>%
#  filter(record_id <= 7713) #filter to last record that would be included
#until_Jun30 <- ref_asymp %>%
#  filter(record_id <= 8332) #filter to last record that would be included
#until_Jul31 <- ref_asymp %>%
#  filter(record_id <= 8959)

until_Jan31 <- until_Jan31 %>% group_by(screen1_assigned, status) %>% summarize(n=n())  
until_Feb28 <- until_Feb28 %>% group_by(screen1_assigned, status) %>% summarize(n=n())  
until_Mar31 <- until_Mar31 %>% group_by(screen1_assigned, status) %>% summarize(n=n())  
until_Apr30 <- until_Apr30 %>% group_by(screen1_assigned, status) %>% summarize(n=n())  
until_May31 <- until_May31 %>% group_by(screen1_assigned, status) %>% summarize(n=n())  
until_Jun30 <- until_Jun30 %>% group_by(screen1_assigned, status) %>% summarize(n=n())  
until_Jul31 <- until_Jul31 %>% group_by(screen1_assigned, status) %>% summarize(n=n())
until_Aug31 <- until_Aug31 %>% group_by(screen1_assigned, status) %>% summarize(n=n())
until_Sep30 <- until_Sep30 %>% group_by(screen1_assigned, status) %>% summarize(n=n())
until_Oct05 <- until_Oct05 %>% group_by(screen1_assigned, status) %>% summarize(n=n())

#verification status

ref_asymp2 = result2 %>% 
  mutate(status2=ifelse((is.na(screen1_veri) & screening_api_complete==1), "Ready to be assigned for verification", NA),
         status2=ifelse((!is.na(screen1_veri) & screening_api_complete==1), "Assigned for verification", status2),
         status2=ifelse((!is.na(screen1_veri) & screening_api_complete==0), "Verified: resolution needed", status2),
         status2=ifelse((!is.na(screen1_veri) & screening_api_complete==2), "Verified: Done", status2))
ref_asymp2 <- ref_asymp2[1874:length(ref_asymp2$record_id),]

until_Jan31_veri <- ref_asymp2 %>%
  filter(record_id <= 5296) #filter to last record that would be included
until_Feb28_veri <- ref_asymp2 %>%
  filter(record_id <= 5856) #filter to last record that would be included
until_Mar31_veri <- ref_asymp2 %>%
  filter(record_id <= 6535) #filter to last record that would be included
until_Apr30_veri <- ref_asymp2 %>%
  filter(record_id <= 7176) #filter to last record that would be included
until_May31_veri <- ref_asymp2 %>%
  filter(record_id <= 7713) #filter to last record that would be included
until_Jun30_veri <- ref_asymp2 %>%
  filter(record_id <= 8332) #filter to last record that would be included
until_Jul31_veri <- ref_asymp2 %>%
  filter(record_id <= 8959)
until_Aug31_veri <- ref_asymp2 %>%
  filter(record_id <= 9483)
until_Sep30_veri <- ref_asymp2 %>%
  filter(record_id <= 10074)
until_Oct05_veri <- ref_asymp2 %>%
  filter(record_id <= 10154)

veri_jan31 <- nrow(until_Jan31_veri[until_Jan31_veri$status2 == "Ready to be assigned for verification" | 
                                    until_Jan31_veri$status2 == "Assigned for verification",])
veri_feb28 <- nrow(until_Feb28_veri[until_Feb28_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Feb28_veri$status2 == "Assigned for verification",])
veri_mar30 <- nrow(until_Mar31_veri[until_Mar31_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Mar31_veri$status2 == "Assigned for verification",])
veri_apr30 <- nrow(until_Apr30_veri[until_Apr30_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Apr30_veri$status2 == "Assigned for verification",])
veri_may30 <- nrow(until_May31_veri[until_May31_veri$status2 == "Ready to be assigned for verification" | 
                                      until_May31_veri$status2 == "Assigned for verification",])
veri_jun30 <- nrow(until_Jun30_veri[until_Jun30_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Jun30_veri$status2 == "Assigned for verification",])
veri_jul31 <- nrow(until_Jul31_veri[until_Jul31_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Jul31_veri$status2 == "Assigned for verification",])
veri_aug31 <- nrow(until_Aug31_veri[until_Aug31_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Aug31_veri$status2 == "Assigned for verification",])
veri_sep30 <- nrow(until_Sep30_veri[until_Sep30_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Sep30_veri$status2 == "Assigned for verification",])
veri_oct05 <- nrow(until_Oct05_veri[until_Oct05_veri$status2 == "Ready to be assigned for verification" | 
                                      until_Oct05_veri$status2 == "Assigned for verification",])

progress$verification = c(veri_jan31, veri_feb28, veri_mar30, veri_apr30, veri_may30, veri_jun30, veri_jul31, veri_aug31, veri_sep30, veri_oct05)

progress$disagreements = c(nrow(until_Jan31_veri[until_Jan31_veri$status2 == "Verified: resolution needed",]), 
                           nrow(until_Feb28_veri[until_Feb28_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Mar31_veri[until_Mar31_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Apr30_veri[until_Apr30_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_May31_veri[until_May31_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Jun30_veri[until_Jun30_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Jul31_veri[until_Jul31_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Aug31_veri[until_Aug31_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Sep30_veri[until_Sep30_veri$status2 == "Verified: resolution needed",]),
                           nrow(until_Oct05_veri[until_Oct05_veri$status2 == "Verified: resolution needed",]))

url <- "https://redcap.ispm.unibe.ch/api/"
token <- "F2725F15FE84D2832E2793BB23B0A62B"
formData2 <- list("token"=token,
                  content='report',
                  format='csv',
                  report_id='163',
                  csvDelimiter='',
                  rawOrLabel='raw',
                  rawOrLabelHeaders='raw',
                  exportCheckboxLabel='false',
                  returnFormat='csv'
)
response3 <- httr::POST(url, body = formData2, encode = "form")
included <- httr::content(response3)
nrow(included)

progress$included <- c(nrow(filter(included, record_id <= 5296)),
                       nrow(filter(included, record_id <= 5856)),
                       nrow(filter(included, record_id <= 6535)),
                       nrow(filter(included, record_id <= 7176)),
                       nrow(filter(included, record_id <= 7713)),
                       nrow(filter(included, record_id <= 8332)),
                       nrow(filter(included, record_id <= 8959)),
                       nrow(filter(included, record_id <= 9483)),
                       nrow(filter(included, record_id <= 10074)),
                       nrow(filter(included, record_id <= 10154)))

ext_jan <- sum(nrow(filter(included, record_id <= 5296 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 5296 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 5296 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_feb <- sum(nrow(filter(included, record_id <= 5856 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 5856 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 5856 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_mar <- sum(nrow(filter(included, record_id <= 6535 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 6535 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 6535 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_apr <- sum(nrow(filter(included, record_id <= 7176 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 7176 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 7176 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_may <- sum(nrow(filter(included, record_id <= 7713 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 7713 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 7713 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_jun <- sum(nrow(filter(included, record_id <= 8332 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 8332 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 8332 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_jul <- sum(nrow(filter(included, record_id <= 8959 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 8959 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 8959 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_aug <- sum(nrow(filter(included, record_id <= 9483 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 9483 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 9483 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_sep <- sum(nrow(filter(included, record_id <= 10074 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 10074 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 10074 & (quest___5 == 1 & question_3_complete != 2)))) 
ext_oct <- sum(nrow(filter(included, record_id <= 10154 & (quest___1 == 1 & question_1_complete != 2))),
               nrow(filter(included, record_id <= 10154 & (quest___2 == 1 & question_2_complete != 2))),
               nrow(filter(included, record_id <= 10154 & (quest___5 == 1 & question_3_complete != 2))))

progress$data_extraction <- c(ext_jan, ext_feb, ext_mar, ext_apr, ext_may, ext_jun, ext_jul, ext_aug, ext_sep, ext_oct)

```

## Screening progress

Since June 2020 we have screened **`r screen_complete`** records out of a total **`r totaltoscreen`** records identified in our search (**`r percent`**).

There are **`r unassigned`** records that have not yet been assigned for screening.

The following reviewer(s) have been assigned a total of **`r assigned_screen`** records:

```{r screen_table, echo=FALSE}
tmp2 %>%
  arrange(status) %>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

```{r screenplot, echo=FALSE}
screenplot
```

## Verification progress

We have verified **`r verified`** records in total.

**`r unassigned_veri`** records are waiting to be assigned for verification.

**`r resolve`** disagreements need to be resolved.

The following reviewer(s) have been assigned a total of **`r assigned_veri`** records:

```{r veri_table, echo=FALSE}
tmpveri %>%
  arrange(status2) %>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

```{r veriplot, echo=FALSE}
veriplot
```

## Overall progress with review

The following table gives an overview of our progress with each month's search results.

The numbers for screening, verification, disagreements, and data extraction indicate the records remaining.

Please note that as records in each stage are complete, more records will move to the next stage, i.e., once a disagreement is resolved, a record will be ready for extraction.

```{r progress_table, echo=FALSE}
progress %>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```